---
title: "CH1 Generalized Linear Models"
author: "Richard Podkolinski"
date: "February 28, 2016"
output:
  knitrBootstrap::bootstrap_document:
    title: "CH1 Generalized Linear Models"
    theme: journal
    highlight: tomorrow
    theme.chooser: TRUE
    highlight.chooser: TRUE
---

```{r Prerequisites, message=FALSE, warning=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(car)
```


# Generalized Linear Models

## 1.9 Data Examples

### 1.9.1 Number of Customers in a "do-it-yourself" store

The variables are as follows:

* $x_1:$ the number of housing units in the district (counting each apartment separately)
* $x_2:$ the average income of people in that district
* $x_3:$ the average age of a housing unit in the district
* $x_4:$ the distance to the nearest competative store
* $x_5:$ the distance to the particular "do-it-yourself" store
* $Y:$ the number of customers of the city that visited the store within those weeks


```{r Do-it-yourself 1, fig.height=15, fig.width=15, bootstrap.thumbnail.size=10}
df = read.table("Data/doityourself.txt", header=TRUE)
glm_data = glm(y ~ x1 + x2 + x3 + x4 + x5, data=df, family = poisson)
summary(glm_data)
par(mfrow=c(2,2))
plot(glm_data)
```

Now with a quasi-likelihood fit

```{r Do-it-yourself 2, fig.height=15, fig.width=15, bootstrap.thumbnail.size=10}
glm_ql = glm(y ~ x1 + x2 + x3 + x4 + x5, data=df, family = quasipoisson)
summary(glm_ql)
par(mfrow=c(2,2))
plot(glm_ql)
```

\textcolor{red}{@TODO: Interpret Results}

### 1.9.2 Buying a New Car?

The variables are as follows:

* $x_1:$ the annual income of a family
* $x_2:$ age of the oldest car the family owned, in years
* $Y:$ whether or not the family purchased a car within a year

Fit an appropriate model to this data. In addition, make a prediciton based on the model that a family with an annual income of $55,000 and a present car of 4 years will buy a new car within a year.


```{r Buying a New Car}
cr = read.table("Data/buyacar.txt", header=TRUE)
cr_m1 = glm(y ~ x1 + x2, data=cr, family=binomial(link = logit))
cr_m2 = glm(y ~ x1 + x2, data=cr, family=binomial(link = probit))
cr_m3 = glm(y ~ x1 + x2, data=cr, family=binomial(link = cloglog))

summary(cr_m1)
summary(cr_m2)
summary(cr_m3)

predict(cr_m1, newdata = data.frame(x1 = 55, x2 = 4))
predict(cr_m1, newdata = data.frame(x1 = 55, x2 = 4), type = "response")
predict(cr_m2, newdata = data.frame(x1 = 55, x2 = 4))
predict(cr_m2, newdata = data.frame(x1 = 55, x2 = 4), type = "response")
predict(cr_m3, newdata = data.frame(x1 = 55, x2 = 4))
predict(cr_m3, newdata = data.frame(x1 = 55, x2 = 4), type = "response")
```

\textcolor{red}{@TODO: Interpret Results}


### Mesquite trees data

We wish to construct a model for the total production fo photosynthetic biomass of mesquite trees by using easily measured aspects of the plant as opposed to actual harvesting of the mesquite.

The variables are as follows:

* $x_1:$ canopy diameter (in meters) measured along hte longest axis of the tree parallel to the ground
* $x_2:$ canopy diameter (in meters) measured along the shortest axis of the tree parallel to the ground
* $x_3:$ total height (in meters) of the tree
* $x_4:$ canopy height (in meters) of the tree
* $y:$ total weight (in grams) of photosynthetic material derived from the actual harvesting of msequite

A multiplicative model is more natural than a linear one as leaf weight should be nearly proportional to canopy volume, and canopy volume should be nearly proportional to the product of canopy dimensions:

$$ Y = \beta_0 x^{\beta_1}_1 x^{\beta_2}_2 x^{\beta_3}_3 x^{\beta_4}_4 \varepsilon $$

A log transformaton gives us a linear model:

$$ {Y}' = {\beta}'_0 + \beta_1 {x_1}' + \beta_2 {x_2}' + \beta_3 {x_3}' + \beta_4 {x_4}' + \varepsilon $$

Where ${\beta}'_0 = \log(\beta_0)$, ${x}'_j = \log(x_j)$ and ${\varepsilon}' = \log(\varepsilon)$.

**Exercise**
Based on the selected output, using $\alpha = 0.05$:

* Perform the model utility test
* Test in the full Model $H_0 : \beta_2 = 0$ vs $H_a: \beta_2 \ne 0$
* Test $H_0: \beta_3 = \beta_4$ vs $H_a: \beta_3 \ne \beta_4$

```{r Mesquite Tree Data}
mq = read.table("Data/mesquite.txt", header = TRUE)
names(mq) = c("x1", "x2", "x3", "x4", "y")

mq_m1 = lm(log(y) ~ log(x1) + log(x2) + log(x3) + log(x4), data=mq)

drop1(mq_m1, ~., test = "F")
summary(mq_m1)

mq_m2 = lm(log(y) ~ log(x1) + log(x2), data=mq)
drop1(mq_m2, ~., test = "F")
summary(mq_m2)
```


## Exercises 

### 6: SLID 

```{r SLID }
sl = SLID
```

