---
title: "Statistical Modeling: Homework 2016"
author: "Richard Podkolinski"
date: "8 April 2016"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(MASS)
library(faraway)
library(dplyr)
library(ggplot2)
library(GGally)
library(rethinking)
df = motorins
```

## Question 1

We first explore a dataset of third part motor insurance claims. The data is the aggregate of insurance claims given specific conditions, this is possible because at the time consumers were given identical risk arguments[1]. So effectively we have sums of accidents of similar insurance policies.

Before we get to modeling, we should take a look at our data. 

There are 8 variabes:

- **Kilometers**: Ordinal variable, amount of kilometers per year driven.
- **Zone**: Nominal variable defining geographical location.
- **Bonus**: Integer, no-claims bonus. Number of years plus one since last claim.
- **Make**: Nominal variable, defining car model, 9 is other.
- **Insured**: Real, number of insured in policy-years.
- **Claims**: Integer, number of claims.
- **Payment**: Integer, Total value of payments in Swedish Korona. 
- **perd**: Real, payment per claim and our response variable.

Even a cursory reading of these variables indicates that there is a going to be a multicollinarity problem in the data. The response variable is a ratio of two of the predictors (**pred** = **Payment** / **Claims**) thus the inclusion of both in any model is not advised. 


We then plot the density of the response variable. 

```{r, echo=FALSE}
df %>% ggplot(aes(x = perd)) + geom_density(color = "slateblue", size = 1) + theme_bw() + ggtitle("Density of Payment per Claim")

## Annotate Outliers 
```

As it is stated in the question, we see a positive value, positively skewed distribution. We also see a small pocket of outliers in the tail of the distribution. A gamma distribution would seem appropriate in this context. We make use of the default link function, the inverse link $$ g(\mu_i) = \frac{1}{\mu_i} $$. 

We first use AIC to select the covariates. However, given that there is clearly two variables where there is multicolinarity, we will refrain from using automatic step function as it will crash glm due to a singular matrix. We can, however, easily loop over the entire set of possible models and retrieve the AIC and BIC. 



```{r, message=FALSE, warning=FALSE}
df = motorins

# Change collapse = "+" in covarm row to alter model type.
covars = names(df)[-8]
respn = "perd ~ "
output = data.frame(Form = character(), AIC = numeric(), BIC = numeric())
for(i in 6:1){
  print(i)
  lcovars = combn(covars, i, simplify = F)  
  
  for(i in 1:length(lcovars)){
    covara = lcovars[[i]]
    covarm = paste(covara, collapse = "+")
    gform = as.formula(paste(respn, covarm))
    print(covarm)
    tryCatch({
      fit =  glm(gform, family = Gamma(link = "inverse"), data = df)
      fitaic = AIC(fit)
      fitbic = BIC(fit)
      inn_out = data.frame(Form = covarm, AIC = fitaic, BIC = fitbic)
      output = rbind(output, inn_out)
    }, error = function(e){})
  }
}

output %>%
  arrange(AIC, BIC) %>%
  head()


# Recheck best models

intr_model = perd ~ Kilometres * Zone * Make * Claims * Payment
main_model = perd ~ Make + Claims + Payment

fiti = glm(intr_model, family = Gamma(link = "inverse"), data = df)
fitm = glm(main_model, family = Gamma(link = "inverse"), data = df)

AIC(fiti)
BIC(fiti)

AIC(fitm)
BIC(fitm)



```



[1]: http://www.statsci.org/data/general/motorins.html    "Motorins"